\chapter{Encryption of the DVB}
There are many parts that are needed to provide DVB with a secure way of 
transmitting streams without facing the risk of content getting stolen. The 
most important parts are:

\begin{itemize}
\item Scrambler - explained in chapter \ref{ch:Scrambling}
\item CAS - explained in section \ref{sec:CAS}
\item Common Interface - explained in section \ref{sec:CI}
\item Descrambler - the inverse of a scrambler.
\end{itemize}

\section{CAS} \label{sec:CAS}
Conditional Access is used to make sure that a user fulfills an amount of 
criteria before being able to view content. A CAS consist of an EMM-generator 
and an ECM-generator in addition to some other parts. We are only interrested 
in the ECMG and EMMG since they are the parts that affect the scrambler. The 
ECM is generated using the CW, while the EMM is generated based on information 
related to the user. That information varies, but might relate to what channels 
the user is allowed to access and when the subscription ends. A TV will not 
broadcast any channels without receiving an EMM telling it to do so.

\Warning[Source]{You need to find more sources than just Patrik}

A simple example is that a user needs to pay for TV-services to be able to access
content. The content provider generates an EMM which tells the smart-card whether
the user is allowed to access the requested material or not. The content provider
also generates an ECM based on the CW, which the smart-card decrypts and passes 
to the descrambler if the EMM allows it.

CA-modules. What are they, how and where are they connected to the system? I 
think that the CAM is where you input the smart-card. 

%Är verkligen den här delen nödvändig? Vad ska stå här? Jag måste fundera på vad 
%jag vill lägga in här, om det ska få vara kvar. Vad är relevant? Vilka 
%distributörer som använder vilket system? Det kan man hitta på google. Vad de 
%olika systemen använder för ECMG och EMMG? Det är konfidentiellt.
\subsection{Standards}\Warning[Remove?]{Or keep?}
Some of the CA systems currently in use are Viaccess, Conax, Irdeto, NDS, Strong 
and NagraVision. You need to buy a specific CAM depending on what distributor 
you choose for your channels. NDS is used by Viasat, Conax is used by Com Hem, Viaccess is used by Boxer and Strong is used by Canal Digital.

\subsubsection{Viaccess}
Viaccess is the CA-module used by Boxer. This is currently the most commonly used
 distributor in Sweden.
%Boxer supports CI+

\subsubsection{Conax}
Conax is the CA-module used by Com Hem. This is currently the second most used
distributor in Sweden. Conax has got its headquarter in Norway and is a part of the Telenor Group, which deals with mobile telecommunications \citep{Conax}.
%Com Hem supports CI+

\subsubsection{Strong}
Strong is the CA-module used by Canal Digital. This is currently the third most 
used distributor in Sweden.
%Canal Digital does not support CI+

\section{DVB-SimulCrypt} \label{sec:Simul}
DVB-SimulCrypt is widespread in Europe, and works as an interface between the 
head-end and the CAS \citep{SimulCrypt:2008}.

SimulCrypt encourages the use of several CAS at once \citep[p. 17]{SimulCrypt:2008}. This is done by sending the same CW to many CAS at the same time, and then 
allowing them to generate an ECM and EMM based on the CW. The multiplexer in the 
head-end then creates TS packets based on those, since the EMMs will determine 
whether the user is allowed access or not. 
%OERHÖRT CHANSAT DETTA JA OKEJ KOLLA ÖVER DET! Allt efter citatet är chansat.
%Why and how can we do that? Is that because the scrambling is the same, where the only thing differentiating the CAS is the encoding of the control word?

\section{Common Interface} \label{sec:CI}
\Warning[Fill]{This section with some relevant information}
THE COMMON INTERFACE CAN'T READ, SO IT WON'T MIND IF I WRITE IN ALL-CAPS ;\_\_\_;
IT ALSO CAN'T UNDERSTAND EMOJIS, WHICH MEANS THAT I CAN INPUT AS MANY STUPID 
THINGS AS I WANT TO. stupid common interface..

\section{Set-up} \label{sec:setup}
The TS is scrambled using a key which is called a \emph{control word} (CW). The 
control word is at least changed every 120th second, but it is not uncommon to 
change the CW every 10 seconds. This means that breaking the key has very little 
effect since it will just be changed in a few seconds even if you manage to 
break it. The control word is generated randomly to make sure that consecutive 
control words are not related to each other. 

The control word is scrambled, then sent to the content provider. The content 
provider then encrypts the CW as an \emph{entitlement control message} (ECM). 
The content provider also generates an \emph{entitlement management message} 
(EMM) which tells the smart-card if the user is allowed access to the data. The 
ECM and EMM are then sent back to the scrambler where it is attached to the 
scrambled TS. This package is then sent to the user, where the ECM, EMM and TS 
are separated. The ECM and EMM are sent to the box containing the smart-card 
for processing, and the TS is sent directly to the descrambler where it waits 
for the CW. The TS is then descrambled using the CW and the pure data is then 
displayed to the user.
